---
title: "Missing Migrant Project"
author: "Quyen Duong, Nhu Casati Nguyen"
semester: "In the beginning of the 2nd semester"
date: "Last edited `r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false 
      smooth_scroll: false
    toc_depth: 2
    number_section: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.align = "center")
```
```{r data, include=FALSE}
# load packages and csv file
library(dplyr)
library(ggplot2)
library(tidyverse)
library(scales)
library(gridExtra)
library(tidyr)
library(readr)
library(knitr)
library(lintr)
library(janitor)
library(Hmisc)
library(patchwork)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)
library(tm)
library(ggridges)
library(ggpol)
library(conflicted)
library(viridis)
library(reshape2)
library(hrbrthemes)
library(ggmap)
library(maps)
library(mapdata)
# library(RgoogleMaps)
# library(sf)
# library(mapview)


# Set conflict preference
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")


# load and read csv files
df <- read_csv("dataset/MissingMigrants-Global-2021-09-05T18-37-06.csv")
head(df)
glimpse(df)
str(df)
```

# Introduction

The Missing Migrants Project tracks deaths of migrants and those who have gone missing along migratory routes across the globe. The research behind the project began following two shipwrecks off the coast of Lampedusa in October 2013, in which nearly 400 migrants lost their lives. Since then, the Missing Migrants Project has developed into an important hub of data and information, which is used widely by media, policy-makers, and researchers. The Missing Migrants Project hosts the only existing global database on migrant deaths.
More than 32,000 migrants around the world have lost their lives since the project began in 2014. More than half of these deaths were recorded in the Mediterranean Sea, which has seen a devastating spike in deaths in recent years. The Missing Migrants Project collaborates with various government and non-government entities to collect data, and also draws on media and other available sources of information.
<https://gmdac.iom.int/missing-migrants-project-0> Footnote^[This is in the footnote.]
Still have to correct here !!!

## Cleaning data

```{r prepare-data}
# convert all variable names to lowercase and replace spaces with underscores
df <- df %>%  
  clean_names() 

# convert reported_date from character to date format 
df <- df %>%  
  mutate(reported_date = parse_date(reported_date, format = "%B %d, %Y"))

# Separate location_coordinates to latitude and longitude 
df <- df %>%
  filter(location_coordinates  != "0,0") %>%
  separate(location_coordinates, into = c("lat", "lon"), sep = ",")
```


## Variable description


```{r summary statistics}
# overview summary statistics of the dataframe
summary(df)

# describe and give summary for each variable
describe(df)

# overview histograms of all numerical variables
df_numerical <- df[, sapply(df, is.numeric)] 
hist.data.frame(df_numerical)

# count missing values and percentage of missing values in each column
d_missing_values <- df %>%
  select(everything()) %>%  
  summarise_all(funs(sum(is.na(.)))) %>%  
  bind_rows(df %>% 
              summarise_each(funs(100 * mean(is.na(.))))) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(number_missing_values = V1, 
         percentage = V2) %>% 
  mutate(percentage = round(percentage, 1))
d_missing_values
```

# Data analysis


```{r set up theme for ggplot}
my_theme_general <- theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        text = element_text(size = 11)) 

my_theme_bar_chart <- theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        text = element_text(size = 11))

my_theme_heat_map <-  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        text = element_text(size = 11),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Mapping incidents

```{r map}
# Select the data dead & missing, lon, lat
d_dead_missing_lon_lat <- df %>% 
  select(total_dead_and_missing, lon, lat) %>% 
  ## to filter wrong values outside of the earth
  filter(as.numeric(lat) >= -80 & as.numeric(lat) <= 75)

# test <- df %>% 
#   select(total_dead_and_missing, lon, lat) %>% 
#   filter(as.numeric(lat) >= -80 & as.numeric(lat) <= 75)
# test2 <- as_tibble(test)
# test3 <- st_as_sf(test2, coords = c("lon", "lat"))
# mapview(test3)


# Map of total dead & missing
world <- map_data("world")

ggplot() + 
  geom_polygon(data = world, aes(x = long, y = lat, group = group), 
               fill = NA, color = "#adb5bd") +
  coord_fixed(1.3) +
  geom_point(data = d_dead_missing_lon_lat,
             aes(x = as.numeric(lon), y = as.numeric(lat), 
                 color = total_dead_and_missing, size = total_dead_and_missing),
             alpha = 0.5) +
  scale_color_gradient(low = "#277da1", high = "#b7094c") +
  labs(col = "Total dead & missing") +
  ggtitle("Total death and missing map") +
  theme(plot.title = element_text(hjust = 0.5)) 

```

## Deaths and dissappearances

```{r total of dead & missing}
# scatter plot number_dead & minimum_estimated_number_of_missing
d_filter_number_dead <- df %>% 
  filter(number_dead <= 400)

ggplot(d_filter_number_dead, aes(number_dead, minimum_estimated_number_of_missing)) + 
  geom_point(alpha = 0.7, color = "#264653" ) +
  labs(title = "Number of dead vs minimun estimated number of missing",
       x = " Number of dead", y = "Minimun estimated number of missing") +
  my_theme_general + 
  theme(panel.grid.minor = element_blank())

# number of dead by region
  d_dead_region <- df %>% 
  count(region, wt = number_dead, sort = TRUE)

## ordered flipped bar chart number of dead by region
ggplot(d_dead_region, 
       aes(x = reorder(region, n), y = n)) +
  geom_col(fill = "#415a77") + 
  geom_text(aes(label = n, y = n + 500), size = 3, hjust = 0) +
  labs(title = "Number of dead by region", x = "Region", y = "Number of dead") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 25000)) +
  my_theme_bar_chart
  
# total of dead & missing by region
d_dead_missing_region <- df %>% 
  count(region, wt = total_dead_and_missing, sort = TRUE)

## ordered flipped bar chart total of dead & missing by region
plot_dead_missing_region <- ggplot(d_dead_missing_region, 
                                   aes(x = reorder(region, n), y = n)) +
  geom_col(fill = "#415a77") +
  geom_text(aes(label = n, y = n + 500), hjust = 0, size = 3) +
  labs(title = "Total of dead and missing by region", 
       x = "Region", y = "Total dead and missing") +
  coord_flip() + 
  scale_y_continuous(limits = c(0, 25000)) +
  my_theme_bar_chart

## way 1: compare the two plots side by side
plot_dead_region + plot_dead_missing_region

## way 2: make both number_dead and total_dead_and_missing in one plot
### create data frame with two columns
d_two_case <- df %>%  
  group_by(region) %>% 
  summarise(sum_dead = sum(number_dead, na.rm = TRUE), 
            sum_total_dead_and_missing = 
              sum(total_dead_and_missing, na.rm = TRUE)) %>% 
  arrange(-sum_dead)
 
### plot two variables in one flip bar chart
ggplot(d_two_case, aes(x = reorder(region, sum_total_dead_and_missing))) +
  geom_point(aes(y= sum_dead), size = 2, color = "red") +
  #geom_text(aes(label = sum_dead), size = 3, hjust = 0) +
  geom_col(aes(y = sum_total_dead_and_missing), alpha = 0.7, fill = "#415a77") + 
  geom_text(aes(label = sum_total_dead_and_missing, y = sum_total_dead_and_missing), 
            size = 3, hjust = -0.6) +
    labs(title = "Comparison dead and total dead and missing by region", 
         subtitle = "Red dot = number of dead, blue bar = Total of dead and missing",
         x = "Region", y = "Number of case") +
    coord_flip() +
    scale_y_continuous(limits = c(0, 25000)) +
    # scale_color_manual(values = c("number_dead" = "red",
    #                               'Total_dead_and_missing' = "#264653")) +
    my_theme_bar_chart

## violin plot total of dead & missing by region
ggplot(df, aes(x = region, y = total_dead_and_missing)) +
  geom_violin(binaxis = "y", 
               stackdir = "center", 
               dotsize = 0.2, 
               inherit.aes = TRUE,
               method = "jitter") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

## strip chart
stripchart(total_dead_and_missing ~ region, data = df, vertical = TRUE,
pch = 16, method = "jitter")

## density plot total dead & missing by region
df %>%  
  ggplot(aes(x = total_dead_and_missing, y = region)) + 
  geom_jitter(alpha = 0.2, width = 0.1) +
    ggridges::geom_density_ridges(color = "#0077b6", fill = "#c3baba") +
  scale_x_continuous(limits = c(0, 100)) +
  labs(title = "Distribution of total dead and missing by region",
       y = "Region", x = "Total dead and missing") +
  my_theme_general +
  theme(plot.title = element_text(hjust = 0.5))

## Jitter plot total dead & missing by region
df %>%  
  ggplot(aes(x = region, y = total_dead_and_missing)) + 
  geom_jitter(alpha = 0.25, width = 0.2, color = "#2a6f97") +
  labs(title = "Distribution of total dead and missing by region",
       x = "Region", y = "Total dead and missing") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


# Total of dead & missing by year
d_dead_missing_year <- df %>% 
  ## exclude 2021 because data do not cover full year
  filter(reported_year != 2021) %>% 
  count(reported_year, wt = total_dead_and_missing, sort = TRUE) 

## Line chart total dead & missing by year
plot_dead_missing_year <- ggplot(d_dead_missing_year, 
                                 aes(x = reported_year, y = n),) +
  geom_line(color = "#415a77", size = 1) +
  geom_point(color = "#415a77", size = 3) +
  geom_text(aes(label = n), size = 3, hjust = -0.7, vjust = 1) +
  labs(title = "Total dead and missing by year", 
       x = "Year", y = "Total dead and missing") +
  scale_x_continuous(breaks = seq(2014, 2021, 1)) +
  scale_y_continuous(limits = c(0, 8500)) +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.minor = element_blank())

## Line chart: comparison total dead & missing in 3 years (2017 - 2020)
df %>% 
  group_by(reported_year, reported_month) %>% 
  filter(reported_year %in% c(2017, 2018, 2019, 2020)) %>%  
  summarise(sum_dead_missing = sum(total_dead_and_missing)) %>% 
  ungroup() %>%
  ggplot(aes(x = factor(reported_month, levels = month.abb), 
             y = sum_dead_missing, 
             color = as.factor(reported_year), 
             group = reported_year)) +
  geom_line(aes(), size = 1) +
  geom_point(size = 2) +
  scale_y_continuous(limits = c(0, 1000)) +
  labs(title = "Comparison on the total dead and missing by month from 2017 to 2020", 
       x = "Month", 
       y = "Total dead & missing") +
  scale_color_manual(values = c("#f72585", "#52b788", "#fe7f2d", "#415a77")) +
  guides(col=guide_legend("Reported year")) +
  my_theme_general

# Total of dead & missing by month
d_dead_missing_month <- df %>% 
  filter(reported_year != 2021) %>% 
  count(reported_month, wt = total_dead_and_missing, sort = TRUE) 


## Lollipop chart total dead & missing by month
ggplot(d_dead_missing_month, 
       aes(x = factor(reported_month, levels = month.abb), y = n)) +
  geom_segment(aes(x = factor(reported_month, levels = month.abb), 
                   xend = yefactor(reported_month, levels = month.abb), 
                   y = 0, yend = n), size = 1.5, color = "#52b788") + 
  geom_point(color = "#415a77", size = 8, alpha = 0.8) +
  geom_text(aes(label = n, y = n), size = 3, vjust = -2.2) +
  labs(title = "Total dead and missing by month", 
       x = "Month", y = "Total dead & missing") +
  scale_y_continuous(limits = c(0, 4100)) +
  my_theme_general + 
  theme(panel.grid.major = element_blank())


# make a new column for season
df <-df %>% 
  mutate(
    season = case_when(
      reported_month %in% c("Oct", "Nov", "Dec") ~ "Fall",
      reported_month %in%  c("Jan", "Feb", "Mar")  ~ "Spring",
      reported_month %in%  c("Apr", "May", "Jun")  ~ "Summer",
      TRUE ~ "Winter"))

# check total dead & missing by season
d_dead_missing_season <- df %>%
  filter(reported_year != 2021) %>% 
  count(season, wt = total_dead_and_missing, sort = TRUE)

## plot with season
df %>%
  ## make an order for seasons
  mutate(ordered_season = fct_relevel(season, c("Spring", "Summer", 
                                                "Fall", "Winter"))) %>%
  ggplot(aes(x = ordered_season, y = total_dead_and_missing)) +
  geom_col()


# Total of dead & missing by migration_route
d_dead_missing_migration_route <- df %>% 
  count(migration_route, wt = total_dead_and_missing, sort = TRUE) %>%
  ## change NA value to "Unknown" to sort the plot
  mutate(migration_route = ifelse(is.na(migration_route), 
                                  "Unknown", migration_route))

## flip bar chart total dead & missing by migration route
plot_dead_missing_migration_route <- ggplot(d_dead_missing_migration_route, 
                                            aes(x = reorder(migration_route, n), 
                                                y = n)) + 
  geom_col(fill = "#415a77") +
  geom_text(aes(label = n), size = 3, hjust = -0.5, vjust = 1) +
  labs(title = "Total dead and missing by migration route", 
       x = "Migration route", y = "Total dead & missing") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 21000)) +
  my_theme_bar_chart

## jitter plot total dead & missing by migration route
df %>%  
  mutate(migration_route = ifelse(is.na(migration_route), 
                                  "Unknown", migration_route)) %>% 
  ggplot(aes(x = migration_route, y = total_dead_and_missing)) + 
  geom_jitter(alpha = 0.35, width = 0.25, color = "#2a6f97") +
  labs(title = "Distribution of total dead and missing by migration route",
       x = "Migration route", y = "Total dead and missing") +
  theme_minimal() +
  # coord_flip() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

## density plot total dead & missing by migration route
df %>%  
  ggplot(aes(x = total_dead_and_missing, y = migration_route)) + 
  geom_jitter(alpha = 0.2, width = 0.1) +
    ggridges::geom_density_ridges(color = "#0077b6", fill = "#c3baba") +
  scale_x_continuous(limits = c(0, 100)) +
  labs(title = "Distribution of total dead and missing by migration route",
       y = "Migration route", x = "Total dead and missing") +
  my_theme_general +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r male-female-children}
# Create a function to calculate sum and percentage of male, female & children
function_gender_children <- function(data) {
  data %>% 
    summarise(sum_men = sum(number_of_males, na.rm = TRUE), 
              sum_women = sum(number_of_females, na.rm = TRUE),
              sum_children = sum(number_of_children, na.rm = TRUE),
              sum_total_dead_and_missing = sum(total_dead_and_missing, na.rm = TRUE),
               ) %>%  
    mutate(perc_known_male = round(100 * (sum_men / sum_total_dead_and_missing), 1), 
           perc_known_female = round(100 * (sum_women / sum_total_dead_and_missing), 1),
           perc_known_children = round(100 * (sum_children / sum_total_dead_and_missing),1))
}

# Overview on percentage of known male, female & children on the total dead & missing
d_perc_known_male_female_children <- df %>%  
  function_gender_children()
d_perc_known_male_female_children

# Percentage of know male, female & children by region
d_perc_known_male_female_children_by_region <- df %>%
  group_by(region) %>% 
  function_gender_children()
d_perc_known_male_female_children_by_region

# percentage male, female, children by region
d_male_female_children_perc_region <- df %>%  
  select(region, number_of_children, number_of_males, number_of_females) %>% 
  group_by(region) %>%  
  summarise(total_children = sum(number_of_children, na.rm = TRUE), 
            total_males = sum(number_of_males, na.rm = TRUE), 
            total_females = sum(number_of_females, na.rm = TRUE)) %>%
  ## pivot long 
  pivot_longer(cols = c(total_children, total_males, total_females)) %>%
  ## calculate the percentage per region -> important
  group_by(region) %>%  
  mutate(perc = 100 * (value / sum(value))) %>% 
  ungroup()

## Stack bar chart male, female, children by region
d_male_female_children_perc_region %>%  
  ## exclude Central Asia due to no data
  filter(region != "Central Asia") %>%  
  ggplot(aes(x = region, y=perc, fill = name)) + 
  scale_y_continuous(labels = percent_format(scale = 1)) +
  geom_col(color = "white") + 
  geom_text(aes(label = ifelse(round(perc) < 3, "", paste0(round(perc), "%"))), 
            position = position_stack(vjust = 0.5), size = 3, color = "white") + 
  scale_fill_manual(values = c("#06d6a0","#f72585", "#355070")) +
  coord_flip() +
  labs(title = "Male, female, children total dead and missing by region", 
       y = "Nomalised percentage of dead & missing",
       x = "Region") + 
  my_theme_bar_chart

# percentage of know male, female & children by month
d_perc_known_male_female_children_by_month <- df %>% 
  group_by(reported_month) %>% 
  function_gender_children()
d_perc_known_male_female_children_by_month

# number of male, female, children by year
d_known_male_female_children_by_year <- df %>% 
  group_by(reported_year) %>% 
  function_gender_children() %>% 
  select(reported_year, sum_men, sum_women, sum_children) %>% 
  filter(reported_year!= 2021) %>% 
  ## important to pivot long here
  pivot_longer(cols = c(sum_men, sum_women, sum_children))



## dodge bar chart comparison male, female, children by year
d_known_male_female_children_by_year %>%  
  mutate(name = fct_relevel(name, c("sum_men, sum_women, sum_children"))) %>% 
ggplot(aes(reported_year, value)) +   
  geom_bar(aes(fill = name), width = 0.6,
           position = position_dodge(0.6), stat="identity") +
  scale_fill_manual(values = c("#06d6a0", "#355070", "#f72585")) +
  labs(title = "Comparison men, women and children by year",
       x = "Year", y = "Total dead and missing") +
  scale_x_continuous(breaks = seq(2014, 2020, 1),
                     labels = seq(2014, 2020, 1)) +
  my_theme_general + 
  theme(panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank())

# percentage male, female, children by year
d_known_male_female_children_by_year <- df %>% 
  group_by(reported_year) %>% 
  function_gender_children()  
  
  # select(reported_year, , sum_women, sum_children) %>% 
  # filter(reported_year %in% c(2014, 2015, 2016, 2017, 2018, 2019, 2020)) %>% 
  # melt(id.vars = "reported_year")

```


## Cause of death


```{r cause-of-death}
# Investigate on the cause of death column
unique(df[c("cause_of_death")])
length(unique(df$cause_of_death))

# View the cause of death in a descending order
d_cause_before_encode <- df %>% 
  count(cause_of_death, sort = TRUE) 


## Wordcloud using wordcloud2 to show the sentences 
wordcloud2(data = d_cause_before_encode, size = 4)


## Wordcloud using wordcloud to show keywords
wordcloud(words = df$cause_of_death, max.words = 300, 
          random.order = FALSE, colors = brewer.pal(8, "Dark2"),
          rot.per=0.35)

# Encode the variable cause of death from 426 unique values to 18 groups
df <- df %>% 
  mutate(cause_encoded =
           case_when(
             str_detect(cause_of_death, "Sexual|sexual|Rape|rape") ~ "Sexual abuse",
             str_detect(cause_of_death, "Sickness|sickness|diabetic|heart attack|
             meningitis|virus|cancer|bleeding|insuline|inhalation|Organ|Coronary|
             Envenomation|Post-partum|Respiratory|Hypoglycemia|Exhaustion|
             Heat stroke") ~ "Health condition",
             str_detect(cause_of_death, "harsh weather|Harsh weather|Harsh conditions|
                        harsh conditions") ~ "Harsh conditions and weather",
             str_detect(cause_of_death, "Unknown|unknown") ~ "Unknown",
             str_detect(cause_of_death, "Starvation|starvation|
                        Malnutrition|malnutrition") ~ "Starvation",
             str_detect(cause_of_death, "dehydration|Dehydration") ~ "Dehydration",
             str_detect(cause_of_death, "Drowning|drowning|Pulmonary|respiratory|
                        lung|bronchial|pneumonia|Pneumonia") ~ "Drowning",
             str_detect(cause_of_death, "hyperthermia|Hyperthermia") ~ "Hyperthermia",
             str_detect(cause_of_death, "hypothermia|Hypothermia") ~ "Hypothermia",
             str_detect(cause_of_death, "asphyxiation|suffocation") ~ "Suffocation",
             str_detect(cause_of_death, "train|bus|vehicle|truck|boat|car|road|
             van|plane|collision|Train|Bus|Vehicle|Truck|Boat|Car|Road|Van|Plane|
             Collision") ~ "Vehicle",
             str_detect(cause_of_death, "murder|stab|shot|violent|blunt force|
             violence|beat-up|fight|murdered|death|Murder|Stab|Shot|Violent|
             Blunt force| Violence|Beat-up|Fight|Murdered|Death|Hanging|Apache|
             mortar|landmine|Gassed") ~ "Murder",
             str_detect(cause_of_death, "crushed to death|crush|Crush|
                        Rockslide") ~ "Crushed",
             str_detect(cause_of_death, "burn|burns|burned|fire|Burn|Burns|
                        Burned|Fire") ~ "Burned",
             str_detect(cause_of_death, "electrocution|Electrocution") 
             ~ "Electrocution",
             str_detect(cause_of_death, "Fall|fall") ~ "Fallen",
             str_detect(cause_of_death, "crocodile|hippopotamus|hippoptamus") 
             ~ "Killed by animals",
             str_detect(cause_of_death, "exposure|Exposure") ~ "Exposure",
             TRUE ~ "Unknown"
           ))

# Check how many times each cause_encoded appears 
d_cause_encoded <- df %>% 
  count(cause_encoded, sort = TRUE) 

# Check total dead & missing by each cause_encoded
d_dead_missing_cause_encoded <- df %>% 
  count(cause_encoded, wt = total_dead_and_missing)

## Bar chart of count of encoded cause of death
ggplot(d_cause_encoded, 
       aes(x = reorder(cause_encoded, n), y = n)) +
  geom_col(fill = "#415a77") +
  geom_text(aes(label = n), size = 3, hjust = -0.5, vjust = 1) +
  labs(title = "Cause of dead", 
       x = "encoded cause of dead", y = "count") +
  coord_flip() +
  my_theme_bar_chart


## Jitter plot total dead & missing by cause of death
ggplot(df, aes(x = cause_encoded, y = total_dead_and_missing)) +
  geom_jitter(alpha = 0.25, width = 0.2, color = "#7209b7") +
  labs(title = "Distribution of total dead and missing by cause of death",
       x = "Encoded cause of death", y = "Total dead and missing") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## Heat map total dead & missing, cause of death, region
ggplot(df, aes(x = region, y = cause_encoded, fill = total_dead_and_missing)) +
  geom_tile(color = "white", lwd = 1, linetype = 1) +
  # scale_fill_continuous(breaks = c(0, 800)) +
  coord_equal() + 
  scale_fill_gradientn(
    breaks = seq(0, 800, 100),
    colours = c("#1b4965", "#1780a1", "#2e6f95", "#455e89", 
                "#723c70", "#892b64", "#a01a58", "#b7094c"),
    values = scales::rescale(seq(0, 800, 100))) +
  labs(title = "Total incidents by cause of deaths and regions", 
       y = "Encoded cause of dead", x = "Region") +
  my_theme_heat_map


## Heat map total dead & missing, cause of death, migration route
ggplot(df, aes(x = migration_route, y = cause_encoded, 
               fill = total_dead_and_missing)) +
  geom_tile(color = "white", lwd = 1, linetype = 1) +
  scale_fill_viridis(discrete=FALSE) +
  coord_equal() +
  my_theme_heat_map


## Heat map total dead & missing, cause of death, year
ggplot(df, aes(x = migration_route, y = factor(reported_month, levels = month.abb),
               fill = total_dead_and_missing)) +
  geom_tile(color = "white", lwd = 1, linetyle = 1) +
  scale_fill_gradient2(breaks = c(0,100, 200, 300, 400, 500, 600, 700, 800),
                       low = "#003f5c", mid = "#003f5c", high = "#bc5090") +
  my_theme_heat_map


###### Create a 6-in-1 plot for top 6 regions and top 6 causes
# Top 6 regions with the highest total dead & missing
d_top_region <- df %>% 
  select(region, total_dead_and_missing) %>% 
  group_by(region) %>% 
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>% 
  top_n(6)

# Get the 6 overall most frequent cause of dead/missins
d_top_deadliest_causes <- df %>% 
  filter(cause_encoded != "Unknown") %>% 
  group_by(cause_encoded) %>%  
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>% 
  arrange(-total) %>% 
  top_n(6)

## first plot -> not good because different causes for each region
df %>%  
  filter(cause_encoded %in% d_top_deadliest_causes$cause_encoded, 
         region %in% d_top_region$region) %>% 
  group_by(region, cause_encoded) %>% 
  summarise(total_dead_missing = sum(total_dead_and_missing, na.rm = TRUE)) %>% 
  ungroup() %>%  
    ggplot(aes(x = cause_encoded, y = total_dead_missing)) +
  geom_col() +
  facet_wrap(~ region) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

## second try without top 6 reasons -> not good because different causes for each region   
df %>%  
  filter(region %in% d_top_region$region) %>% 
  group_by(region, cause_encoded) %>% 
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>% 
  ungroup() %>%  
    ggplot(aes(x = cause_encoded, y = total)) +
  geom_col() +
  facet_wrap(~ region) +
  coord_flip()


#### Each region has their different top reason, we have to plot each region
# Mediterranean top 5 cause of death
plot1 <- df %>%
  filter(region %in% "Mediterranean") %>%
  group_by(cause_encoded) %>%
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>%
  top_n(n = 5) %>%
  arrange(desc(total)) %>% 
  ggplot( aes(x = reorder(cause_encoded, total), y = total) ) +
  geom_bar(stat="identity", fill="steelblue", alpha=.6, width=.4) +
  ylab("Total dead & missing") +
  xlab("Mediterranean") +
  coord_flip() + 
  my_theme_bar_chart

# North Africa top 5 cause of death
plot2 <- df %>%
  filter(region %in% "North Africa") %>%
  group_by(cause_encoded) %>%
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>%
  top_n(n = 5) %>%
  arrange(desc(total)) %>% 
  ggplot( aes(x = reorder(cause_encoded, total), y = total) ) +
  geom_bar(stat="identity", fill="steelblue", alpha=.6, width=.4) +
  ylab("Total dead & missing") +
  xlab("North Africa") +
  coord_flip() + 
  my_theme_bar_chart

# US-Mexico Border top 5 cause of death
plot3 <- df %>%
  filter(region %in% "US-Mexico Border") %>%
  group_by(cause_encoded) %>%
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>%
  top_n(n = 5) %>%
  arrange(desc(total)) %>% 
  ggplot( aes(x = reorder(cause_encoded, total), y = total) ) +
  geom_bar(stat="identity", fill="steelblue", alpha=.6, width=.4) +
  ylab("Total dead & missing") +
  xlab("US-Mexico Border") +
  coord_flip() + 
  my_theme_bar_chart

# Sub-Saharan Africa top 5 cause of death
plot4 <- df %>%
  filter(region %in% "Sub-Saharan Africa") %>%
  group_by(cause_encoded) %>%
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>%
  top_n(n = 5) %>%
  arrange(desc(total)) %>% 
  ggplot( aes(x = reorder(cause_encoded, total), y = total) ) +
  geom_bar(stat="identity", fill="steelblue", alpha=.6, width=.4) +
  ylab("Total dead & missing") +
  xlab("Sub-Saharan Africa") +
  coord_flip() + 
  my_theme_bar_chart

# Southeast Asia top 5 cause of death
plot5 <- df %>%
  filter(region %in% "Southeast Asia") %>%
  group_by(cause_encoded) %>%
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>%
  top_n(n = 5) %>%
  arrange(desc(total)) %>% 
  ggplot( aes(x = reorder(cause_encoded, total), y = total) ) +
  geom_bar(stat="identity", fill="steelblue", alpha=.6, width=.4) +
  ylab("Total dead & missing") +
  xlab("Southeast Asia") +
  coord_flip() + 
  my_theme_bar_chart

# Horn of Africa top 5 cause of death
plot6 <- df %>%
  filter(region %in% "Horn of Africa") %>%
  group_by(cause_encoded) %>%
  summarise(total = sum(total_dead_and_missing, na.rm = TRUE)) %>%
  top_n(n = 5) %>%
  arrange(desc(total)) %>% 
  ggplot( aes(x = reorder(cause_encoded, total), y = total) ) +
  geom_bar(stat="identity", fill="steelblue", alpha=.6, width=.4) +
  ylab("Total dead & missing") +
  xlab("Horn of Africa") +
  coord_flip() + 
  my_theme_bar_chart

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol = 2, nrow = 3)


```



## Survivors

```{r survivors}
# create a new variable: total death, missing & survivors 
df <- df %>% 
  mutate(total_dead_missing_survivor = 
           ifelse(is.na(total_dead_and_missing), 
                  0, 
                  total_dead_and_missing) + 
           ifelse(is.na(number_of_survivors), 
                  0, 
                  number_of_survivors))


# percentage of known survivors on the total death, missing & survivors
df %>% 
  summarise(perc_survivor = round(100 * (number_of_survivors / 
                                           total_dead_missing_survivor), 1))

# df %>%
#    replace(is.na(.), 0) %>%
#    mutate(sum = rowSums(total_dead__missing, number_of_survivors))

# ranking on the number of survivors by region
d_survivor_region <- df %>% 
  count(region, wt = number_of_survivors, sort = TRUE) 

## lolipop chart survivor by region
ggplot(d_survivor_region, aes(x = reorder(region, n), y = n)) +
  geom_segment(aes(x = reorder(region, n), xend = reorder(region, n), 
                   y = 0, yend = n), size = 1, color = "skyblue") + 
  geom_point( color="blue", size = 4, alpha = 0.6) +
  geom_text(aes(label = n, y = n + 500), size = 3, hjust = 0) +
  labs(title = "Number of survivor by region", 
       x = "Region", y = "Number of survivor") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 50000)) +
  my_theme_bar_chart

# ranking on the number of survivors by year
df %>% 
  count(reported_year, wt = number_of_survivors, sort = TRUE) 

# ranking on the number of survivors by month
df %>% 
  count(reported_month, wt = number_of_survivors, sort = TRUE) 

# ranking on the number of survivors by migration_route
d_survivor_migration_route <- df %>% 
  count(migration_route, wt = number_of_survivors, sort = TRUE) %>% 
  ## change NA value to "Unknown" to sort the plot
  mutate(migration_route = ifelse(is.na(migration_route), 
                                  "Unknown", migration_route))

## lolipop chart survivor by region
ggplot(d_survivor_migration_route, aes(x = reorder(migration_route, n), y = n)) +
  geom_segment(aes(x = reorder(migration_route, n), 
                   xend = reorder(migration_route, n), 
                   y = 0, yend = n), size = 1, color = "skyblue") + 
  geom_point( color="blue", size = 5, alpha = 0.6) +
  geom_text(aes(label = n, y = n + 500), size = 3, hjust = -0.5) +
  labs(title = "Number of survivor by migration route", 
       x = "Migration route", y = "Number of survivor") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 45000)) +
  my_theme_bar_chart
```

## Source

```{r source-quality}
## Heat map source by region filled by total incidents
df_source <- df %>%
  group_by(region, source_quality) %>%
  summarise(total_incidents = sum(total_dead_and_missing))

ggplot(df_source, aes(region, source_quality, fill= total_incidents)) + 
  geom_tile(color = "white", lwd = 1, linetype = 1) +
  scale_fill_viridis(discrete=FALSE) +
  coord_equal() +
  xlab("Region") +
  ylab("Source quality") +
  ggtitle("Source quality and region") +
  my_theme_heat_map

## Dodge bar chart source by region
df %>% 
  group_by(region, source_quality) %>% 
  filter(!(region %in% c("Northern Africa", "East Asia", "Central Asia", 
                         "North America", "Eastern Africa"))) %>% 
  count(source_quality) %>% 
  ggplot(aes(x = region, y = n, 
             fill = as.factor(source_quality))) +
  geom_col() +
  coord_flip() +
  my_theme_general
  

# Calculate source proportion in each region
df %>% 
  group_by(region, source_quality) %>% 
  ## filter the regions with too few data points
  filter(!(region %in% c("Northern Africa", "East Asia", "Central Asia",
                         "North America", "Eastern Africa"))) %>%
  count(source_quality) %>%
  group_by(region) %>% 
  mutate(perc = 100 * (n / sum(n))) %>% 
  ungroup() %>%
  ## Stack bar chart
  ggplot(aes(x = region, y = perc, fill = as.factor(source_quality))) + 
  scale_y_continuous(labels = percent_format(scale = 1)) +
  geom_col(color = "white") + 
  geom_text(aes(label = ifelse(round(perc) <3, "", paste0(round(perc), "%"))),
            position = position_stack(vjust = 0.5), 
            size = 3, color = "white") +
  # scale_fill_manual(values = c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")) +
  scale_fill_manual(values = c("#c3e7ff","#96bed9", 
                               "#6a96b3", "#3e708f", "#004c6d")) +
  coord_flip() +
  labs(title = "Source quality by region", 
       y = "Nomalised percentage of source quality",
       x = "Region") + 
  guides(fill=guide_legend(title="Source quality")) +
  my_theme_bar_chart
  

```


```{r model}
# create df for the model, here we do not use dplyr to match with the package
df_red <- df[c("total_dead_and_missing","lat","lon")]
df_red <- as.data.frame(sapply(df_red, as.numeric))
colnames(df_red) <- c("z", "y", "x")
df_red <- df_red[c("x", "y", "z")]

library(MASS)
library(spatial)

# Fits a trend Surface by Least-squares 
surf <- surf.ls(6, df_red)
# Evaluate trend surface over a grid
trsurf <- trmat(surf, min(df_red$x), max(df_red$x),
                min(df_red$y), max(df_red$y), 50)
# Regression diagnostics for trend surfaces
infl_surf <- trls.influence(surf)
cand <- as.data.frame(infl_surf)[abs(infl_surf$stresid) > 1.5,]
cand_xy <- df_red[as.integer(rownames(cand)), c("x", "y")]

eqscplot(trsurf, type = "n")
contour(trsurf, add = TRUE, col = "grey")
plot(surf, add = TRUE, div = 100)
points(cand_xy, pch = 16, col = "orange")
title(main = "Surface mapping")
```




